### 2020-10-17 readme.Rmd
some package cleaning
cleaning up package

[![Travis build status](https://travis-ci.com/skiptoniam/ppmData.svg?branch=master)](https://travis-ci.com/skiptoniam/ppmData)

[![Codecov test coverage](https://codecov.io/gh/skiptoniam/ppmData/branch/master/graph/badge.svg)](https://codecov.io/gh/skiptoniam/ppmData?branch=master)

## ppmData is an R package that can be used to set up a quadrature scheme for multiple species point processes modelling and could be used in [ecomix](https://github.com/skiptoniam/ecomix).

The package aims to setup a model matrix for point process modelling for multiple species (marked) point processes. The package is based on the quadrature scheme proposed by Berman & Turner 1993, and Warton & Shepard 2010. The idea behind this package is simple, provided a set of presences for a species (or multiple species) and a spatial region, we can generate generate a model matrix which can be included in a Poisson point process model.  The code does provide the ability to produce a background point data-set without the inclusion of presences (a dummy grid) or for single species (which was developed as a proof of concept for the multispecies approach).

We also provide the capacity to generate quasi-random points are an alternative to grid-based or random background point designs. Quasi-random (sampling) background points are an advanced form of spatially-balanced survey design or point stratification, that aims to reduce the frequency of placing samples close to each other (relative to simple randomization or grid designs). A quasi-random background point design improves efficiency of background point sampling (and subsequent modelling) by reducing the amount of spatial auto-correlation between data implying that each sample is providing as much unique information as possible (Grafston & Tille, 2013) and thus reducing low errors for geostatistical prediction (Diggle & Ribeiro, 2007).

For a worked example with single and multiple species presences see the package vignette.

>>>>

[![Travis build status](https://travis-ci.org/skiptoniam/ppmData.svg?branch=master)](https://travis-ci.org/skiptoniam/ppmData)

## ppmData is an R package that can be used to set up a quadrature scheme for multiple species point processes modelling and could be used in [ecomix](https://github.com/skiptoniam/ecomix) or for other multiple species Point Processes.

The approach uses quasi-random sampling to generate a quadrature
scheme based (e.g Berman & Turner 1992; Foster et al, 2017). Quasi-random 
sampling quadrature are form of spatially-balanced survey design or 
point stratification, that aims to reduce the frequency of placing samples 
close to each other (relative to pseudo-random or grid designs). A 
quasi-random quadrature design improves efficiency of background point 
sampling (and subsequent modelling) by reducing the amount of spatial 
auto-correlation between data implying that each sample is providing as much 
unique information as possible (Grafston & Tille, 2013) and thus reducing low 
errors for geostatistical prediction (Diggle & Ribeiro, 2007).
The weights each quasi-random point in the quadrature scheme is calculated 
using a dirichlet tesselation (Turner 2020). To improve comptational efficiency 
\link[deldir]{deldir} funcion for a large number of of quadrature points, we 
set up an approach which breaks up the problem into manageable sub-windows. 
We do this by keeping each deldir call to less that 5000 points (which appears 
to be the point where the algorithm slowes noticibly). To avoid edge effect 
(large areas on the edges of sub-areas), we rotate the subregions three times, 
the first two use a set of prime number to closest to total number of points 
(presences+quadrature points) divided by 5000. This allows us to split the study
region in to an subsetable number of subwindows. We then take the median weight 
across all weight calculated for each point.
